---
title: 'Statistical report: Covid-19 vs RAASi'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data sources

### SCB

Län (Stockholm), civil status and family type are from RTB as of 2019-12-31. 
Age, sex, country of birth and number of children is also from 2019-12-31.
Disposable income and education are from LISA 2018.

### Socialstyrelsen

Comorbidities and covid-19 outcome are derived from the National Patient Registry 
(in-patient or out-patient) from 1997 (1998 Skåne) 
and treatments from the Dispensed Drug Registry. 

Note that data after 2018 is preliminary for the National Patient Registry and 
Cause of Death Registry. From Socialstyrelsens homepage:
"Patientregistret kommer göras tillgängligt för utlämnanden för forskningsprojekt
och statistikbeställningar redan en eller två månader efter referensmånaden. De
data som görs tillgängliga kommer delvis vara ofullständiga, framför allt
kommer det med hög sannolikhet att saknas diagnoskoder för delar av
materialet. Regionala skillnader avseende kvalitet kommer därför att finnas i de
data som görs tillgängliga. Fullständiga data i patientregistret beräknas finnas tre
månader efter rapporterad period.
Dödsorsaksintyg för dödsfall som är relaterade till covid-19 kommer kodas med
förtur, så att data om sådana dödsfall kommer vara tillgängliga cirka fyra veckor
efter dödsfallet"

Utöver detta information från handläggare från Socialstyrelsn e-post: 
"patientregistret slutenvård år 2020 kan det finnas pågående vårdtillfällen 
(dvs. vårdtillfällen utan  utskrivningsdatum). Jag vet inte hur ni hanterar 
dessa, men det kan vara bra för dig att veta. Eftersom diagnos ofta fastställt 
vid utskrivning kan diagnoserna på dessa vårdtillfällen vara lite ”skakiga”. 

## Time period

Index date is defined as `r global_indexdtm` (first covid-19 case in Sweden end Jan). 
End of follow-up is defined as `r global_enddtm`.

## Population 

The population consists of a person with any of the following: hypertension, 
diabetes, ischemic heart disease, heart failure, kidney disease defined 
from the NPR (in-patient or out-patient) from 1997 (1998 Skåne) 
up until 2020-02-01 and alive and not migrated at 2020-02-01. 

```{r flow}
mykable(flow, caption = "Flowchart")
```

\clearpage
\newpage

## Created variables 

```{r outcodes}
mykable(metaout, caption = "Comorbidities and outcomes from NPR")
```

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

### Outcome Covid-19

The outcome is defined as hospitalized (in-patient) confirmed covid-19 (U071) as main 
diagnosis (HDIA) in the National Patient Registry OR as confirmed covid-19 (U071)
as underlying cause of death (ULORSAK) in Cause of Death Registry. 

```{r, child = "../src/deskstatcovid.Rmd"}

```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Some missing data for scb variables (not sex, age, number of children, region). 
In models including these variables observations with missing values are excluded. 

## Propensity scores

A propensity score for treatment with RASi and MRA respectively was estimated for 
each patient with logistic regression with variables indicated in Table \ref{tab:tab1}. 
Continuous variables were modeled using 
cubic splines with 4 dfs. The propensity score was then incorporated in the model 
by inverse probability weights together with the original variables. 
Weights were truncated at 10.  

## Baseline characteristics

```{r, child = "../src/venn.Rmd"}

```

```{r, child = "../src/tab1.Rmd"}

```

```{r, child = "../src/tab1covid.Rmd"}

```

## Covid-19

The association between RASi/MRA and Covid-19 was investigated using logistic 
regression models for partly the whole cohort and partly predefined subgroups. 

```{r, child = "../src/outcovidtab.Rmd"}

```

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor for Adj ps IPW model in the whole cohort. 

\clearpage
\newpage

### Unmeasured confounders

Interpretation is that if there is an unmeasured confounder with an association 
with covid as 1.3 (harmful), for example crappy eGFR, and crappy eGFR is present 
in the RASi group with 40% and in the no RASi group with 60%, the effect of RASi 
would be non-significant. If you up the OR to 2 (similar proportions) you have 
a harmful effect and if you up the OR to 3 and you have 10% of this confounder 
in the RASi and 80% in the no RASi you will have a VERY harmful effect. [@unmeasuredc]

```{r, child = "../src/forest_unmeasuredconf.Rmd"}

```

\clearpage
\newpage

## All-cause mortality after Covid-19

The association between RASi/MRA and all-cause mortality after Covid-19 was 
investigated using cox regression models for partly the whole cohort and 
partly predefined subgroups. Only patients with Covid-19 were included in the 
models.

```{r, child = "../src/deskstatdeaths.Rmd"}

```

NOTE: Socialstyrelsen are prioritizing coding Covid-19 deaths. 
That means the probably a larger proportion of the uncoded deaths are not 
covid-19 compared to the coded.   

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the Adj ps IPW model in the 
whole cohort. Possible outliers were visually inspected by plotting the dfbetas. 

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

`r sessionInfo()$R.version$version.string` [@r] is used, for more details see below.

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/covid19rasi. On publication 
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
