---
title: 'Statistical report: Covid-19 vs RASi/MRA'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data sources

### SCB

Län (Stockholm), civil status and family type are from RTB as of 2019-12-31. 
Age, sex, country of birth and number of children is also from 2019-12-31.
Disposable income and education are from LISA 2018. Age is from 2018 in the 
negative control analyses (see below). 

### Socialstyrelsen

Comorbidities and covid-19 outcome are derived from the National Patient Registry 
(in-patient or out-patient) from 1997 (1998 Skåne) 
and treatments from the Dispensed Drug Registry. 

Note that data after 2018 is preliminary for the National Patient Registry and 
Cause of Death Registry. From Socialstyrelsens homepage:
"Patientregistret kommer göras tillgängligt för utlämnanden för forskningsprojekt
och statistikbeställningar redan en eller två månader efter referensmånaden. De
data som görs tillgängliga kommer delvis vara ofullständiga, framför allt
kommer det med hög sannolikhet att saknas diagnoskoder för delar av
materialet. Regionala skillnader avseende kvalitet kommer därför att finnas i de
data som görs tillgängliga. Fullständiga data i patientregistret beräknas finnas tre
månader efter rapporterad period.
Dödsorsaksintyg för dödsfall som är relaterade till covid-19 kommer kodas med
förtur, så att data om sådana dödsfall kommer vara tillgängliga cirka fyra veckor
efter dödsfallet"

Information from handläggare at Socialstyrelsen via e-mail: 
"patientregistret slutenvård år 2020 kan det finnas pågående vårdtillfällen 
(dvs. vårdtillfällen utan  utskrivningsdatum). Jag vet inte hur ni hanterar 
dessa, men det kan vara bra för dig att veta. Eftersom diagnos ofta fastställt 
vid utskrivning kan diagnoserna på dessa vårdtillfällen vara lite ”skakiga”. 

## Time period

Index date is defined as `r global_indexdtm` (first covid-19 case in Sweden end Jan). 
End of follow-up is defined as `r global_enddtm`, 
decision based on visual inspection of patterns in NPR and Cause of Death Registry data.

For the negative control analyses the indexdate is instead one year prior `r global_indexdtm - 365` and 
end of follow-up is defined as `r global_enddtm - 365`.  

## Population 

The population recieved from Socialstyrelsens consisted of persons with any of the following: 
hypertension, diabetes, ischemic heart disease, heart failure, kidney disease, obesity or af 
(ICD codes in Table \ref{tab:outcomcodes}) defined 
from the NPR (in-patient or out-patient) from 1997 (1998 Skåne) and alive 2020-01-01. 

```{r flow}
mykable(flow, caption = "Flowchart")
```

The population for the negative control analyses is a subset of the main population but 
patients are requred to have a prior diagnosis of hypertension, diabetes, hf, kidney, 
ihd, being >= 18 years and not have ARNI 
(defined as one dispensed prescripton within 5 months) at `r global_indexdtm - 365` 
and immigrated prior to 2014. 

The median age (IQR) is `r pop %>% summarise(med = dF(median(scb_age), dig = 1),
                                             q1 = dF(quantile(scb_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(scb_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pop %>% count(scb_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(scb_sex == "Female") %>%
  pull(perc)`% females.                                   

\clearpage
\newpage

## Created variables 

```{r outcomcodes}
mykable(metaout, caption = "Comorbidities and outcomes from NPR")
```

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

Note that the comorbidities, treatments and outcomes used in the analyses 
of covid and the negative control analyses are constructed based on different 
time periods.

### Outcome Covid-19

The outcome is defined as hospitalized (in-patient) confirmed covid-19 (U071) as main 
diagnosis (HDIA) in the National Patient Registry OR as confirmed covid-19 (U071)
as underlying cause of death (ULORSAK) in Cause of Death Registry. 

```{r, child = "../src/deskstatcovid.Rmd"}

```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

Some missing data for scb variables (not sex, age, number of children, region). 
In models including these variables observations with missing values are excluded. 

## Propensity scores

A propensity score for treatment with RASi and MRA respectively was estimated for 
each patient with logistic regression with variables indicated in Table \ref{tab:tab1all}. 
Continuous variables (age) were modeled using 
cubic splines with 4 dfs. The propensity score was then incorporated in the model 
using inverse probabilty of treatment weighting, truncated at 10, together 
with the original variables. Sandwich variance estimators were applied 
(prob more appropriate with bootstrap ci but will take too long). 

This was done in a similar manner for the population used for the negative control 
analyses including the variables used for this analyses. 

## Baseline characteristics

```{r, child = "../src/venn.Rmd"}

```

```{r, child = "../src/tab1.Rmd"}

```

## Covid-19

The association between RASi/MRA and Covid-19 was investigated using logistic 
regression models for partly the whole cohort and partly predefined subgroups.
Variables indicated in Table \ref{tab:tab1all} are included in the models where 
continuous variables (age) were modeled using cubic splines with 4 dfs. 

### Assumptions

Outliers were investigated with Cook's distance and multicollinearity 
with the variance inflation factor for Adj IPW model in the whole cohort. 
=> No action deemed necessary. 

\clearpage
\newpage

```{r, child = "../src/outcovidtab.Rmd"}

```

\clearpage
\newpage

### Unmeasured confounders

Unmeasured confounders were assessed according to [@unmeasuredc]. 

```{r, child = "../src/forest_unmeasuredconf.Rmd"}

```

All unmeasured confounders are presumed to be present 
in the RASi/MRA group with `r p[2] * 100`% and in the no RASi/MRA group 
with `r p[1] * 100`%. 
Interpretation: if there is an unmeasured confounder with an association 
with covid as `r orucrasi[1]` (harmful), for example eGFR, and eGFR is present 
in the RASi group with `r p[2]*100`% and in the no RASi group with `r p[1]*100`%, the effect of RASi 
would be non-significant. If you up the OR to `r orucrasi[2]` for the unmeasured confounder 
you have a statistically significant harmful effect and if you up the OR to `r orucrasi[3]` 
very harmful effect (>1.5). 

\clearpage
\newpage

## All-cause mortality after Covid-19

```{r, child = "../src/deskstatdeaths.Rmd"}

```

The association between RASi/MRA and all-cause mortality after Covid-19 was 
investigated using cox regression models for partly the whole cohort and 
partly predefined subgroups. Only patients with Covid-19 were included in the 
models. Variables indicated in Table \ref{tab:tab1all} are included in the models 
where continuous variables (age) were modeled using cubic splines with 4 dfs.

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the Adj IPW model in the 
whole cohort. Possible outliers were visually inspected by plotting the dfbetas. 
=> Some signs of non-proportial hazards for RASi but no action deemed necessary. 

\clearpage
\newpage

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/outtab.Rmd"}

```

### Unmeasured confounders

Unmeasured confounders were similar assessed according to [@unmeasuredc]. 

```{r, child = "../src/forest_unmeasuredconf_death.Rmd"}

```

All unmeasured confounders are presumed to be present 
in the RASi/MRA group with `r p[2] * 100`% and in the no RASi/MRA group 
with `r p[1] * 100`%. 

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

`r sessionInfo()$R.version$version.string` [@r] is used, for more details see below.

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/covid19raasi. On publication 
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
