---
title: 'Statistical report: Covid-19 vs RAASi'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

source("./src/vars.R")

cacheon <- TRUE
```             

# Data handling

## Data source

XXX

Note that data after 2018 is preliinary and XXXXXX.  

## Index date

Index date is defined as 2020-02-01 (first covid-19 case in Sweden end Jan).

## Population 

The population consists of a person with any of the following: hypertension, diabetes, ischemic 
heart disease, heart failure, kidney disease defined from the patientregistry 
(in-patient or out-patient) from 1997 (1998 Skåne) up until 2020-02-01 and 
alive and not migrated at 2020-02-01. 

```{r flow}
mykable(flow, caption = "Flowchart")
```

## Created variables 


```{r outcodes}
mykable(outcommeta, caption = "Comorbidities and outcomes from NPR")
```

Outcomes covid and negative controls are after index date (2020-02-01). 
Other outcomes are after first confirmed covid.  

Note to self: Länka sv besök för utfall???

Note to self: Rerun prep_sosdata (remove missing INDATUM)

The treatments were derived from the Dispensed Drug Registry where 
use was defined as dispensed drug within 5 months prior to index date. 
The following ATC codes were used to define the respective treatments: 

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Will be some missing data for education but otherwise pretty much no missing data. 
Imputation is hassle (and can limit chooice of models) so try to avoid. 

## Propensity scores

A propensity score for treatment with raas/arni, arni (within subgroup with HF) 
and mra respectivly was estimated for each patient with 
logistic regression with variables indicated in Table \ref{tab:tab1}. 
Continuous variables were modeled using 
cubic splines with 4 dfs. The propensity score was then incorporated in the model 
by inverse probability weighted together with the original variables. 
Weights were truncated at 10.  

Note to self: bootstrapped ci??? computational time = :(.

## Baseline characteristics

```{r, child = "../src/venndiagram.Rmd"}

```

```{r, child = "../src/tab1.Rmd"}

```

## Outcomes

The following outcomes are considered: 

1. confirmed covid-19 (U071) in inpatient PAR HDIA or DORS ULORSAK
2. covid-19 (U071, U072, Z861A, KVÅ ZV100) in PAR (in-and out-patient, any position) 
or DORS any position

The primary endpoint is considered to be confirmed covid-19. 

### Subgroups

By indication. 

Only people in Stockholms län. 

```{r, child = "../src/outcovidtab.Rmd"}

```


```{r, child = "../src/outtab.Rmd"}

```

```{r, child = "../src/km.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

`r sessionInfo()$R.version$version.string` [@r] is used, for more details see below.

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: XXX. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
